### 2013-07-29 paper.tex
Hell Week Monday 11AM update.

We study a class of maps between Bott-Samelson bimodules.  
% These maps are represented by certain types of diagrams, as described in Section~\ref{sec:prelim_map}.  The maps are composed by juxtaposition.  
Bott-Samelson bimodules arise from the study of Soergel bimodules, as mentioned in \cite{gr4all}, in the sense that a description of the category of Bott-Samelson bimodules is sufficient to understand that of the Soergel bimodules. 
However, Bott-Samelson bimodules admit a combinatorial description.  In particular, \cite{gr4all} gives a presentation for the category of Bott-Samelson bimodules in terms of planar diagrammatics (i.e. diagrams in the plane).

@@

Given a sequence consisting of the letters $s$ and $t$ of length $n$, as well as a binary string of length $n$, one can construct such a diagram using an algorithm known as Libedinsky's light leaves, as detailed in Section~\ref{sec:prelim_lightleave}.  
These light leaves are interesting because they form a spanning set for the maps we are interest in, as shown in \cite{gr4all}.
Since two diagrams may be composed, we may construct a new map given two binary strings by composing the results of two applications of the light leaves.  The goal of the project is to describe the relationship between the final result of such a composition and the two binary strings.

The technical description of the problem statement is quite involved.  Thusly, the preliminaries for the problem are given in Appendix~\ref{sec:prelim_background}, while the definitions used in this paper are recorded in Section~\ref{sec:prelim}.  The problem statement is given in Section~\ref{sec:probstate}.
The results begin by solving the special case of one color completely in Section~\ref{sec:res_onecolor}.  In Section~\ref{sec:res_count_maps}, we also present a formula for the number of maps between two given bimodules.  The most interesting results are in Sections~\ref{sec:res_alt_restrict} and \ref{sec:res_alt_compute}, where we investigate the case where the given string is $ststst\cdots$.  In particular, we show that the possible resulting structures is extremely limited, and provide a recursion based on these restrictions for special cases.

@@

In general, a string of $s$'s and $t$'s will be referred to as an \emph{expression}.  In the case where there are no consecutive identical characters (e.g. the string is either $\AA_n$ or $\AA_n'$), we will refer to them as \emph{reduced expressions}.

@@

Finally, given a string $\ul x$ (either a binary string or a string of $s$'s and $t$'s) we will denote the $i\th$ character as $\ul x[i]$.  In general, underline Roman letters denote strings.

\subsubsection{Definitions for Maps and Diagram}
Consider a string $\ul r$ of letters $s$ and $t$.  Given a binary string $\ul b$, we will let $\MM_{\ul r}(\ul b)$ denote the map (or its associated diagram; we will refer to maps and their associated diagrams interchangeably) formed by the light leaves, as prescribed in Appendix~\ref{sec:prelim_lightleave}.  We will refer to such structures as \emph{half-maps} or \emph{half-diagrams} should such a need arise.  Additionally, let $\Top_{\ul r}(\ul b)$ denote the top of the half-diagram $\MM_{\ul r}(\ul b)$.

@@

\begin{definition*}
	A connected component which (i) is acyclic, and (ii) does not touch the top or bottom boundaries is called a \emph{barbell}.
\end{definition*}
Notice that, by homotopy, every blue barbell is simply $\barbell{barbell_blue}$, which we will see later evaluates as $\alpha_s$.  Similarly, every red barbell is simply $\barbell{barbell_red} = \alpha_t$.

\begin{definition*}
	A \emph{fence} is a contiguous path which runs from the top of the boundary to the bottom of the boundary (i.e. paths between the labelled vertices).  The diagram is divided by these fences into \emph{pastures}, which will be numbered from left to right as $0$, $1$, \dots, $k$, where $k$ is the number of fences.  
\end{definition*}
\begin{definition*}
	A component is called \emph{attached} if it is connected to a fence.  Otherwise, it is \emph{free}.  By convention, any attached components belong to the pasture to the left of the associated fence.
\end{definition*}
An example of a fence complete with two barbells is given in Figure~\ref{fig:def_barbell_fence}.

@@

\begin{definition*}
	A \emph{bubble} is a bounded face alongside with any components contained within it.  The \emph{color} of the bubble is the color of the edges which form the bounded face.  
\end{definition*}
\begin{definition*}
	We say that the bubble $B$ \emph{holds} a component $c$ if $c$ is inside $B$, but there is no other bubble $B'$ inside $B$ which contains $c$.  
\end{definition*}

@@

\begin{definition*}
	A \emph{caterpillar} consists of a free component and the contents of any bubbles it contains.  A caterpillar \emph{holds} a component if some bubble of the caterpillar holds it.
\end{definition*}
An example of a caterpillar is given in Figure~\ref{fig:def_caterpillar}.  Notice that any barbell or bubble is a special case of a caterpillar.

@@

In order to study these maps as algebraic structures, we wish to consider full-maps as polynomials in $\RR[x,y][\alpha_s,\alpha_t]$.  Hence, we establish the following conventions:
\begin{enumerate}
	\ii The map $x \mapsto fx$ will be abbreviated as $f$ from this point forward.  
	\par Note that, in particular, $\barbell{barbell_blue} = \alpha_s$.  Furthermore, one can check that, say, $\barbell{barbell_blue}\barbell{barbell_blue}\barbell{barbell_red}\barbell{alpha_blue} = \alpha_s^2 \alpha_t$.  
	Please note that $x \mapsto xf$ does not get a similar abbreviation.  So, $\barbell{alpha_blue}\barbell{barbell_blue} \neq \alpha_s$.
	\ii We will work \emph{modulo lower terms}.  That is, any map which contains a connected component that touches any boundary only once (total) is considered zero.  So, for example, we have $\barbell{break_blue} = 0$.
\end{enumerate}
Henceforth, we will refer to diagrams and their associated polynomials interchangeably.

One can verify that, using the barbell-forcing rules, we can recursively move all barbells to the left, and any ``broken'' walls are reduced to zero, or are contracted.  Hence, any full-diagram can be reduced to linear combinations of maps of the form $f \mapsto \alpha_s^m \alpha_t^n f$.  In other words, every full diagram can be represented as a polynomial in $\RR[x,y][\alpha_s,\alpha_t]$.

In fact, we can even show that, with this new convention, that the following identity holds for every $f$:
\begin{equation}
	\barbell{alpha_blue} f = \partial_s(f) \barbell{break_blue} + s(f) \barbell{alpha_blue}
	\label{eq:barbell_break}
\end{equation}
where $\partial_s$ is the Demazure operator defined in Section~\ref{sec:prelim_genrel}.

\subsubsection{Quantum Numbers}
A particularly nice sequence of polynomials involved can be defined recursively.  They are termed \emph{two-color quantum numbers}.
\begin{definition*}[Quantum Numbers]
	For each integer $n \ge 0$, define two polynomials $[n]_x, [n]_y \in \ZZ[x,y]$ recursively by $[0]_x = [0]_y = 0$, $[1]_x = [1]_y = 1$, $[2]_x = x$, $[2]_y = y$, and
	\begin{align*}
		[2]_x[n]_y = [n+1]_x - [n-1]_x \\
		[2]_y[n]_x = [n+1]_y - [n-1]_y
	\end{align*}
	Finally, $[-n]_x = -[n]_x$ and $[-n]_y = -[n]_y$ for each $n \ge 0$.
\end{definition*}
The quantities $[n]_x$ and $[n]_y$ are closely related.  In particular, $[2k+1]_x = [2k+1]_y$ and $[2]_x[2k]_y = [2]_y[2k]_x$ for all integers $k$.

The quantum numbers are relevant to the problem because of the identity
\begin{equation}
	s([n]_x\alpha_s + [n-1]_y\alpha_t) = [n]_x\alpha_s + [n+1]_y\alpha_t
	\label{eq:quantum}
\end{equation}
and its analogous form for $t$.  One can verify \eqref{eq:quantum} immediately follows from definitions.  This allows us to recursively compute the polynomial which results from pushing a barbell through multiple alternating fences.  For example, $ststst(\alpha_s) = ststst([1]_x\alpha_s + [0]_y\alpha_t) = [7]_x\alpha_s + [6]_y\alpha_t$.

@@

It is easy to observe that if $\ul r$ is a reduced expression of length $n+1$, then \[ \MM_{\ul r} ( \underbrace{11\dots11}_{\text{$n$ 1's}}0, \underbrace{11\dots11}_{\text{$n$ 1's}}0 ) = Q_{\ul r}. \]

@@

The motivation for this problem arises from the following definition.
\begin{definition*}
	A map is \emph{idempotent} if multiplying it with itself returns the original graph.
\end{definition*}
Colloquially, $m^2=m$ if and only if $m$ is idempotent.

Fix an $\ul r$.  If we can find two maps $A$ and $B$ for which $AB$ is the identity, then $(BA)^2=BABA=BA$ will be an idempotent.  This motivates us to construct potential idempotents by taking two light leaves whose top sequences are identical, and composing $A$ upside-down with $B$.  If $AB=1$, then this will yield an idempotent.

@@

\section{A Complete Characterization of the One-Color Case}
\label{sec:res_onecolor}
In this section we provide a complete characterization for the $\MM_{\SS_n}(\ul a, \ul b)$.  % The proof is effectively trivial, but requires some rather irritating details and definitions in order to be clear.

First, we need a criteria to determine the top.  Fortunately, this is extremely easy.
\begin{definition*}
	For a binary string $\ul b$, let $\pi_1(\ul b)$ denote the number of $1$'s in $\ul b$, and $\pi_0(\ul b)$ the number of $0$'s.
\end{definition*}
\begin{proposition}
	For any $\ul b \in \BB_n$, \[
		\Top_{\SS_n}(\ul b) =
		\begin{cases}
			\varnothing & \text{if } \pi_1(\ul b) \equiv 0 \pmod{2} \\
			s & \text{if } \pi_1(\ul b) \equiv 1 \pmod{2}.
		\end{cases}
		\]
	\label{prop:s_top}
\end{proposition}
\begin{proof}
	Straightforward induction on $n$.  
\end{proof}

Subsequently, we make the following definition for this section only.
\begin{definition*}
	For a binary string $\ul b$, define the \emph{partial-sum string} of $\ul b$, denoted $\ul b^\ast$, as follows:
	\[
		\ul b^\ast [i] = 
		\begin{cases}
			1 & \text{if } \ul b[1] + \ul b[2] + \dots + \ul b[i] \equiv 1 \pmod{2} \\
			0 & \text{otherwise}.
		\end{cases}
	\]
\end{definition*}
This lets us state our main result for this section. 
\begin{theorem}
	If $\ul a, \ul b \in \BB_n$ obey $\Top_{\SS_n}(\ul a) = \Top_{\SS_n}(\ul b)$, then
	\[
		\MM_{\SS_n} \left( \ul a, \ul b \right)
		=
		\begin{cases}
			0 & \text{if }\ \exists 1 \le i \le n-1:\ \ul a^\ast[i] = \ul b^\ast[i] = 1 \\
			\alpha_s^{n - \pi_1(\ul a^\ast) - \pi_1(\ul b^\ast) - \ul a^\ast[n]} & \text{otherwise}.
		\end{cases}
	\]
	\label{thm:s_eval}
\end{theorem}
The proof is essentially only a matter of finding loops and counting connected components.
As such, it is deferred to in Appendix~\ref{sec:s_proof};

\section{Counting Maps}
\label{sec:res_count_maps}
We give a formula for the number of binary strings $\ul b$ such that $\Top_{\ul r}(\ul b) = \ul T$ for each expression $\ul T$ and $\ul r$.  One can check that, by construction, $\Top(\ul b)$ is always a reduced expression.

@@

\begin{proof}[Proof of Proposition \ref{thm:count_maps}]
	First we will resolve the case where $\ul r = \AA_n$, so that $\ell = n$ and $2^{\ell-n} = 1$.  This follows by induction on $n$.  The main idea is that, for the binary strings of length $n$, we consider those that end with $1$ and those that end with $0$.  Those that end with a $0$ have the same top when the trailing $0$ is deleted, whilst those with end bit $1$ have a top which differs by one character if the trailing $1$ is deleted.  We simply check that in all eight cases, the binomial coefficients add up to give the new binomial coefficients (via Pascal's identity).  The computation is left to a reader.

	To finish to all cases, we note that we can treat each consecutive block of $s$'s of length $m$ as a single $s$, where there a $2^{m-1}$ ways to get a $1$ and $2^{m-1}$ ways to get a $0$.  (This follows from Proposition \ref{prop:s_top}.)  This reduces to the previous case $\ul r = \AA_n$, but with some extra factors of $2$; aggregating all of them yields the factor of $2^{\ell-n}$, as desired.
\end{proof}

@@

We now present three results which show that the possible outputs of a map with base $\AA_n$ are extremely restricted.  
\begin{lemma}[Bubble Lemma]
	Consider a bubble in a full-diagram with base $\AA_n$.  The bubble holds exactly one caterpillar, whose color is opposite that of the bubble.
	\label{thm:bubble}
\end{lemma}
\begin{lemma}[Caterpillar Lemma]
	Any caterpillar in a full-diagram with base $\AA_n$ evaluates in the form $x^my^n\alpha_s$ if it is blue, or $x^my^n\alpha_t$ is red.
	\label{thm:caterpillar}
\end{lemma}
\begin{lemma}[Pasture Lemma]
	Consider a full-diagram with $N+1$ pastures (numbered $0$, $1$, \dots, $N$) with base $\AA_n$ and top $T$.  Suppose further that $N \ge 1$ (or equivalently, $T \neq \varnothing$).  Then:
	\begin{enumerate}[(i)]
		\ii Pastures $1$ through $N-1$ contain either nothing or an attached component.
		\ii Pasture $N$ contains nothing if $T[N] = \AA_n[n]$, and contains a single caterpillar otherwise, with color corresponding to $\AA_n[n]$.
	\end{enumerate}
	\label{thm:pasture}
\end{lemma}
The Bubble Lemma is the key combinatorial observation, which hinges on the fact that any two anchors with the same color have a third anchor of the other color between them.  The other two lemmas follow almost directly from it.
The details of the proofs are given in Appendix~\ref{sec:lemmata_proof}.

>>>>

We study a class of maps between Bott-Samelson bimodules.  
Bott-Samelson bimodules arise from the study of Soergel bimodules, as mentioned in \cite{gr4all}, in the sense that a description of the category of maps between Bott-Samelson bimodules is sufficient to understand the corresponding category for the Soergel bimodules. 
However, the morphisms of Bott-Samelson bimodules admit a combinatorial description.  In particular, \cite{gr4all} gives a presentation for the category of these morphisms in terms of planar diagrammatics (i.e. diagrams in the plane).

@@

Given a sequence consisting of the letters $s$ and $t$ of length $n$, as well as a binary string of length $n$, one can construct such a diagram using an algorithm known as Libedinsky's light leaves, as detailed in Appendix~\ref{sec:prelim_lightleave}.  
These light leaves are interesting because they form a spanning set for the maps we are interested in, as shown in \cite{gr4all}.
The project deals with composing two maps formed in such a manner.
The goal of the project is to describe the relationship between such a composition and the two original binary strings.

The technical description of the problem statement is quite involved.  Thusly, the preliminaries for the problem are given in Appendix~\ref{sec:prelim_background}, while the definitions used in this paper are recorded in Section~\ref{sec:prelim}.  The problem statement is given in Section~\ref{sec:probstate}.
The results begin by solving the special case of one color completely in Section~\ref{sec:res_onecolor}.  In Section~\ref{sec:res_count_maps}, we also present a formula for the number of maps between two given bimodules.  The most interesting results are in Sections~\ref{sec:res_alt_restrict} and \ref{sec:res_alt_compute}, where we investigate the case where the given string is $ststst\cdots$.  In particular, we show that the possible resulting structures are extremely limited, and provide a recursion based on these restrictions for that case.

@@

In general, a string of $s$'s and $t$'s will be referred to as an \emph{expression}.  In the case where there are no consecutive identical characters (in other words, the string is either $\AA_n$ or $\AA_n'$), we will refer to them as \emph{reduced expressions}.

@@

Finally, given a string $\ul x$ (either a binary string or an expression) we will denote the $i\th$ character as $\ul x[i]$.  In general, underlined Roman letters denote strings.

\subsubsection{Definitions for Maps and Diagram}
Consider an expression $\ul r$.  Given a binary string $\ul b$, we will let $\MM_{\ul r}(\ul b)$ denote the map (or its associated diagram; we will refer to maps and their associated diagrams interchangeably) formed by the light leaves, as prescribed in Appendix~\ref{sec:prelim_lightleave}.  We will refer to such structures as \emph{half-maps} or \emph{half-diagrams} should such a need arise.  Figure~\ref{fig:lightleaf_example} (in Appendix~\ref{sec:prelim_lightleave}) gives an example of such a half-map.

Additionally, let $\Top_{\ul r}(\ul b)$ denote the \emph{top} of the half-diagram $\MM_{\ul r}(\ul b)$, as described in Appendix~\ref{sec:prelim_lightleave}.  
Note that the top of any half-map is always a reduced expression.  In fact, one can verify that the top of a map is given explicitly by
\begin{equation}
	\Top_{\ul r}(\ul b) = \prod_{i=1}^n \left( \ul r[i] \right)^{\ul b[i]} \in W.
	\label{eq:top}
\end{equation}

@@

\begin{definition*}
	A connected component which (i) is acyclic, and (ii) does not touch the top or bottom boundaries is called a \emph{barbell}.
\end{definition*}
Notice that, by combining homotopy and contraction (see Appendix~\ref{sec:prelim_genrel}), every blue barbell is simply $\barbell{barbell_blue}$.  Similarly, every red barbell is simply $\barbell{barbell_red}$.

\begin{definition*}
	A \emph{fence} is a contiguous path which runs from the top of the boundary to the bottom of the boundary (i.e. paths between the labelled vertices).  The diagram is divided by these fences into \emph{pastures}.
\end{definition*}
\begin{definition*}
	A component is called \emph{attached} if it is connected to a fence.  Otherwise, it is \emph{free}.  By convention, any attached components belong to the pasture to the left of the fence to which it is attached.
\end{definition*}
An example of a fence complete with two barbells is given in Figure~\ref{fig:def_barbell_fence}.

@@

\begin{definition*}
	A \emph{bubble} is a bounded face alongside with any components contained within the face.  The \emph{color} of the bubble is the color of the edges which form the bounded face.  
\end{definition*}
\begin{definition*}
	We say that the bubble $B$ \emph{holds} a component $c$ if $c$ is inside $B$, but there is no other bubble $B'$ inside $B$ which contains $c$.  
\end{definition*}

@@

\begin{definition*}
	A \emph{caterpillar} consists of a free component and the contents of any bubbles formed by its edges.  A caterpillar \emph{holds} a component if some bubble of the caterpillar holds it.
\end{definition*}
An example of a caterpillar is given in Figure~\ref{fig:def_caterpillar}.  Notice that any barbell or bubble is a special case of a caterpillar.

@@

In order to study these maps as algebraic structures, we wish to consider full-maps as polynomials in $\RR[x,y][\alpha_s,\alpha_t]$.  Hence, we establish the following conventions:
\begin{enumerate}
	\ii The map $x \mapsto fx$ will be abbreviated as $f$ from this point forward.  
	\par Note that, in particular, $\barbell{barbell_blue} = \alpha_s$ follows from a straightforward computation.  Furthermore, one can check that, say, $\barbell{barbell_blue}\barbell{barbell_blue}\barbell{barbell_red}\barbell{alpha_blue} = \alpha_s^2 \alpha_t$.  
	Please note that $x \mapsto xf$ does not get a similar abbreviation.  So, $\barbell{alpha_blue}\barbell{barbell_blue} \neq \alpha_s$.
	\ii We will work \emph{modulo lower terms}.  That is, any map which contains a connected component that touches any boundary only once (total) is considered zero.  So, for example, we have $\barbell{break_blue} = 0$.
\end{enumerate}
Henceforth, we will refer to diagrams and their associated polynomials interchangeably.

One can verify that, using the barbell-forcing rules, we can recursively move all barbells to the left, and any ``broken'' walls are reduced to zero, or are contracted.  Hence, any full-diagram can be reduced to linear combinations of maps of the form $f \mapsto \alpha_s^m \alpha_t^n f$.  In other words, every full-diagram can be represented as a polynomial in $\RR[x,y][\alpha_s,\alpha_t]$.

In fact, we can even show that, with this new convention, that the following identity holds for every $f$:
\begin{equation}
	\barbell{alpha_blue} f = \partial_s(f) \barbell{break_blue} + s(f) \barbell{alpha_blue}
	\label{eq:barbell_break}
\end{equation}
where $\partial_s$ is the Demazure operator defined in Appendix~\ref{sec:prelim_genrel}.

For reference, Appendix~\ref{sec:ststs_matrix} contains a complete table of all maps of the form $\MM_{\AA_5} (\ul a, \ul b)$.

\subsubsection{Quantum Numbers}
\label{sec:quantum}
A particularly nice sequence of polynomials involved can be defined recursively.  They are termed \emph{two-color quantum numbers}.
\begin{definition*}[Quantum Numbers]
	For each integer $n \ge 0$, define two polynomials $[n]_x, [n]_y \in \ZZ[x,y]$ recursively by $[0]_x = [0]_y = 0$, $[1]_x = [1]_y = 1$, $[2]_x = x$, $[2]_y = y$, and
	\begin{align*}
		[2]_x[n]_y = [n+1]_x - [n-1]_x \\
		[2]_y[n]_x = [n+1]_y - [n-1]_y
	\end{align*}
	Finally, $[-n]_x = -[n]_x$ and $[-n]_y = -[n]_y$ for each $n \ge 0$.
\end{definition*}
The quantities $[n]_x$ and $[n]_y$ are closely related.  In particular, $[2k+1]_x = [2k+1]_y$ and $[2]_x[2k]_y = [2]_y[2k]_x$ for all integers $k$.  This follows immediately by induction.

The quantum numbers are relevant to the problem because of the identity
\begin{equation}
	s([n]_x\alpha_s + [n-1]_y\alpha_t) = [n]_x\alpha_s + [n+1]_y\alpha_t
	\label{eq:quantum}
\end{equation}
and its analogous form for $t$.  One can verify \eqref{eq:quantum} immediately follows from definitions.  This gives us a robust way to express the results of action on $\alpha_s$ or $\alpha_t$ by elements of $W$.  For example, $ststst(\alpha_s) = ststst([1]_x\alpha_s + [0]_y\alpha_t) = [7]_x\alpha_s + [6]_y\alpha_t$.

@@

The significance is that, if $\ul r$ is a reduced expression of length $n+1$, then \[ \MM_{\ul r} ( \underbrace{11\dots11}_{\text{$n$ 1's}}0, \underbrace{11\dots11}_{\text{$n$ 1's}}0 ) = Q_{\ul r}. \]
An illustration of such a full-map is given in Figure~\ref{fig:push_quantum}.
This follows from \eqref{eq:barbell_break}.

		picture one;
		draw(one, (4,0)--(4,h), s);
		dot(one, (4,h), dot_s);
		draw(one,(0,0)--(0,3*h), s);
		draw(one,(1,0)--(1,3*h), t);
		draw(one,(2,0)--(2,3*h), s);
		draw(one,(3,0)--(3,3*h), t);

		picture two;
		draw(two, (4,0)--(4,h), s);
		dot(two, (4,h), dot_s);
		draw(two,(0,0)--(0,3*h), s);
		draw(two,(1,0)--(1,3*h), t);
		draw(two,(2,0)--(2,3*h), s);
		draw(two,(3,0)--(3,3*h), t);

		add(one); add(reflect((0,0),(1,0))*two);
		draw((-1,0)--(5,0));
	\end{asy}
	\caption{$\MM(11110,11110)$.}
	\label{fig:push_quantum}
\end{figure}

@@

The motivation for this problem arises from the following definition.
\begin{definition*}
	A map is \emph{idempotent} if multiplying it with itself returns the original map.
\end{definition*}
Colloquially, $m^2=m$ if and only if $m$ is idempotent.

Fix an $\ul r$.  If we can find two maps $A$ and $B$ for which $AB$ is the identity, then $(BA)^2=BABA=BA$ will be an idempotent.  This motivates us to construct potential idempotents by taking two light leaves whose top sequences are identical, and composing $A$ upside-down with $B$, in the hope that $AB=1$.

@@

\section{A Complete Characterization of the One-Color Case}
\label{sec:res_onecolor}
In this section we provide a complete characterization for $\MM_{\SS_n}(\ul a, \ul b)$.  % The proof is effectively trivial, but requires some rather irritating details and definitions in order to be clear.

First, we need a criteria to determine the top.  Fortunately, this is extremely easy.
\begin{definition*}
	For a binary string $\ul b$, let $\pi_1(\ul b)$ denote the number of $1$'s in $\ul b$.
\end{definition*}
\begin{proposition}
	For any $\ul b \in \BB_n$, \[
		\Top_{\SS_n}(\ul b) =
		\begin{cases}
			\varnothing & \text{if } \pi_1(\ul b) \equiv 0 \pmod{2} \\
			s & \text{if } \pi_1(\ul b) \equiv 1 \pmod{2}.
		\end{cases}
		\]
	\label{prop:s_top}
\end{proposition}
\begin{proof}
	Straightforward induction on $n$.  
\end{proof}

Subsequently, we make the following definition for this section only.
\begin{definition*}
	For a binary string $\ul b$, define the \emph{partial-sum string} of $\ul b$, denoted $\ul b^\ast$, as follows:
	\[
		\ul b^\ast [i] = 
		\begin{cases}
			1 & \text{if } \ul b[1] + \ul b[2] + \dots + \ul b[i] \equiv 1 \pmod{2} \\
			0 & \text{otherwise}.
		\end{cases}
	\]
\end{definition*}
This lets us state our main result for this section. 
\begin{theorem}
	If $\ul a, \ul b \in \BB_n$ obey $\Top_{\SS_n}(\ul a) = \Top_{\SS_n}(\ul b)$, then
	\[
		\MM_{\SS_n} \left( \ul a, \ul b \right)
		=
		\begin{cases}
			0 & \text{if }\ \exists 1 \le i \le n-1:\ \ul a^\ast[i] = \ul b^\ast[i] = 1 \\
			\alpha_s^{n - \pi_1(\ul a^\ast) - \pi_1(\ul b^\ast) + \ul a^\ast[n]} & \text{otherwise}.
		\end{cases}
	\]
	\label{thm:s_eval}
\end{theorem}
The proof is essentially only a matter of finding loops and counting connected components.
As such, it is deferred to Appendix~\ref{sec:s_proof}.

\section{Counting Maps}
\label{sec:res_count_maps}
We give a formula for the number of binary strings $\ul b$ such that $\Top_{\ul r}(\ul b) = \ul T$ for each expression $\ul T$ and $\ul r$.

@@

\begin{proof}[Proof of Proposition \ref{thm:count_maps}]
	First we will resolve the case where $\ul r = \AA_n$, so that $\ell = n$ and $2^{\ell-n} = 1$.  This follows by induction on $n$.  The main idea is that, for the binary strings of length $n$ giving a top $T$, we consider those that end with $1$ and those that end with $0$.  Using \eqref{eq:top}, we may then write the new quantity in terms of the number of strings of length $n-1$ with a different top.
	% After deleting the last bit, the former type have a top which differs by one character according to $\AA_n[n]$, while the latter have the same top.
	We simply check that in all eight cases, the binomial coefficients add up to give the new binomial coefficients (via Pascal's identity).  The computation is left to a diligent reader.

	To finish the case $\ell \neq n$, we note that we can treat each consecutive block of $s$'s of length $m$ as a single $s$, where there a $2^{m-1}$ ways to get a $1$ and $2^{m-1}$ ways to get a $0$.  (This follows from Proposition \ref{prop:s_top}.)  This reduces to the previous case $\ul r = \AA_n$, but with some extra factors of $2$; aggregating all of them yields the factor of $2^{\ell-n}$, as desired.
\end{proof}

@@

We now present three results which show that the possible outputs of a map with base $\AA_n$ are extremely restricted.  
\begin{lemma}[Bubble Lemma]
	Consider a bubble in a full-diagram with base $\AA_n$.  The bubble holds exactly one caterpillar, whose color is opposite that of the bubble.
	\label{thm:bubble}
\end{lemma}
\begin{lemma}[Caterpillar Lemma]
	Any caterpillar in a full-diagram with base $\AA_n$ evaluates in the form $x^my^n\alpha_s$ if it is blue, or $x^my^n\alpha_t$ if it is red.
	\label{thm:caterpillar}
\end{lemma}
\begin{lemma}[Pasture Lemma]
	Consider a full-diagram with $N+1$ pastures (numbered $0$, $1$, \dots, $N$) with base $\AA_n$ and top $T$.  Suppose further that $N \ge 1$ (or equivalently, $T \neq \varnothing$).  Then:
	\begin{enumerate}[(i)]
		\ii Pastures $1$ through $N-1$ contain either nothing or an attached component.
		\ii Pasture $N$ contains nothing if $T[N] = \AA_n[n]$, and contains a single caterpillar otherwise.  The caterpillar is blue if $\AA_n[n]$ is $s$, and red otherwise.
	\end{enumerate}
	\label{thm:pasture}
\end{lemma}

Included in Figure~\ref{fig:large_alt_example} is an example of a map with base $\AA_{27}$.   The Bubble Lemma and Pasture Lemma can be readily observed in this example.

\begin{figure}[ht]
	\centering
	\begin{asy}
		size(17cm);
		real h = 0.7;
		pen s = blue, t = red + dashed + 0.6;
		pen dot_s = blue, dot_t = red;
		int n = 27;

		picture one;
		draw(one, (0,0)--(0,h/2)..((0+2)/2.0,h*2)..(2,h/2)--(2,0), s);
		draw(one, (5,0)--(5,h/2)..((5+7)/2.0,h*2)..(7,h/2)--(7,0), t);
		draw(one, (7,0)--(7,h/2)..((7+9)/2.0,h*2)..(9,h/2)--(9,0), t);
		draw(one, (12,0)--(12,h/2)..((12+14)/2.0,h*2)..(14,h/2)--(14,0), s);
		draw(one, (11,0)--(11,h/2)..((11+15)/2.0,h*4)..(15,h/2)--(15,0), t);
		draw(one, (17,0)--(17,h/2)..((17+19)/2.0,h*2)..(19,h/2)--(19,0), t);
		draw(one, (21,0)--(21,h/2)..((21+23)/2.0,h*2)..(23,h/2)--(23,0), t);
		draw(one, (23,0)--(23,h/2)..((23+25)/2.0,h*2)..(25,h/2)--(25,0), t);
		draw(one, (20,0)--(20,h/2)..((20+26)/2.0,h*6)..(26,h/2)--(26,0), s);
		draw(one, (1,0)--(1,h), t);
		dot(one, (1,h), dot_t);
		draw(one, (3,0)--(3,h), t);
		dot(one, (3,h), dot_t);
		draw(one, (6,0)--(6,h), s);
		dot(one, (6,h), dot_s);
		draw(one, (8,0)--(8,h), s);
		dot(one, (8,h), dot_s);
		draw(one, (13,0)--(13,h), t);
		dot(one, (13,h), dot_t);
		draw(one, (18,0)--(18,h), s);
		dot(one, (18,h), dot_s);
		draw(one, (22,0)--(22,h), s);
		dot(one, (22,h), dot_s);
		draw(one, (24,0)--(24,h), s);
		dot(one, (24,h), dot_s);
		draw(one,(4,0)--(4,7*h), s);
		draw(one,(9,0)--(9,7*h), t);
		draw(one,(10,0)--(10,7*h), s);
		draw(one,(15,0)--(15,7*h), t);
		draw(one,(16,0)--(16,7*h), s);
		draw(one,(19,0)--(19,7*h), t);
		dot(one, (7, h/2), dot_t);
		dot(one, (23, h/2), dot_t);
		dot(one, (9, h/2), dot_t);
		dot(one, (15, h/2), dot_t);
		dot(one, (19, h/2), dot_t);

		picture two;
		draw(two, (5,0)--(5,h/2)..((5+7)/2.0,h*2)..(7,h/2)--(7,0), t);
		draw(two, (7,0)--(7,h/2)..((7+9)/2.0,h*2)..(9,h/2)--(9,0), t);
		draw(two, (12,0)--(12,h/2)..((12+14)/2.0,h*2)..(14,h/2)--(14,0), s);
		draw(two, (14,0)--(14,h/2)..((14+16)/2.0,h*2)..(16,h/2)--(16,0), s);
		draw(two, (17,0)--(17,h/2)..((17+19)/2.0,h*2)..(19,h/2)--(19,0), t);
		draw(two, (21,0)--(21,h/2)..((21+23)/2.0,h*2)..(23,h/2)--(23,0), t);
		draw(two, (23,0)--(23,h/2)..((23+25)/2.0,h*2)..(25,h/2)--(25,0), t);
		draw(two, (20,0)--(20,h/2)..((20+26)/2.0,h*6)..(26,h/2)--(26,0), s);
		draw(two, (0,0)--(0,h), s);
		dot(two, (0,h), dot_s);
		draw(two, (1,0)--(1,h), t);
		dot(two, (1,h), dot_t);
		draw(two, (2,0)--(2,h), s);
		dot(two, (2,h), dot_s);
		draw(two, (3,0)--(3,h), t);
		dot(two, (3,h), dot_t);
		draw(two, (6,0)--(6,h), s);
		dot(two, (6,h), dot_s);
		draw(two, (8,0)--(8,h), s);
		dot(two, (8,h), dot_s);
		draw(two, (13,0)--(13,h), t);
		dot(two, (13,h), dot_t);
		draw(two, (15,0)--(15,h), t);
		dot(two, (15,h), dot_t);
		draw(two, (18,0)--(18,h), s);
		dot(two, (18,h), dot_s);
		draw(two, (22,0)--(22,h), s);
		dot(two, (22,h), dot_s);
		draw(two, (24,0)--(24,h), s);
		dot(two, (24,h), dot_s);
		draw(two,(4,0)--(4,7*h), s);
		draw(two,(9,0)--(9,7*h), t);
		draw(two,(10,0)--(10,7*h), s);
		draw(two,(11,0)--(11,7*h), t);
		draw(two,(16,0)--(16,7*h), s);
		draw(two,(19,0)--(19,7*h), t);
		dot(two, (7, h/2), dot_t);
		dot(two, (14, h/2), dot_s);
		dot(two, (23, h/2), dot_t);
		dot(two, (9, h/2), dot_t);
		dot(two, (16, h/2), dot_s);
		dot(two, (19, h/2), dot_t);

		add(one); add(reflect((0,0),(1,0))*two);
		draw((-1,0)--(27,0));
	\end{asy}
	\caption{A large map with base $\AA_{27}$.}
	\label{fig:large_alt_example}
\end{figure}

The Bubble Lemma is the key combinatorial observation, which hinges on the fact that any two anchors with the same color have a third anchor of the other color between them.  The other two lemmas follow almost directly from it.
The details of the proofs are given in Appendix~\ref{sec:lemmata_proof}.
